EDIpack: A generic and interoperable Lanczos-based solver for quantum impurity problems

### 2025-05-30 05_examples.tex
Updates from Overleaf

The system is defined on a Bethe lattice with density of states
$\rho(\e)=\frac{1}{2D}\sqrt{D^2-\e^2}$ This density of states does not normalize to 1,
where $D=2t$ is the half-bandwidth.
In DMFT, this lattice model is mapped onto a quantum impurity problem with
an effective electronic bath that must be determined
self-consistently.

Below, we present the key components of a basic DMFT implementation 
using \NAME for the Bethe lattice, employing either the Fortran or 
C++ APIs. Fully operational versions of the two codes can be found 
in the {\tt examples} directory of the \NAME source code. Both 
samples share a similar initial structure, including memory 
allocation, Bethe lattice creation, and solver initialization. 'Bethe lattice creation' sounds a bit weird here. DMFT does not directly work with a lattice as a collection of sites, it only deals with the spectral functions.

@@

\item[{\tiny {\it User}}] Update the Weiss field via the {\bf self-consistency}
  relation: $\GG^{-1}_0(i\omega) = G_\mathrm{loc}(i\omega) +
    \Sigma(i\omega)$ $G_\mathrm{loc}$ inverted?. For the Bethe lattice, this simplifies to
    $\Delta = \tfrac{D}{4}G_\mathrm{loc}$ Missing square in the coefficient?.

@@

\subsubsection{Finite temperature solution (w2dynamics interface, CT-HYB benchmark)}\label{SecExamplesBetheDMFTw2d}
Capitalization in 'w2dynamics' is different compared to Sec. 4.
To demonstrate the usage of the w2dynamics interface to \NAME,  here we briefly discuss how to solve the same problem, i.e. the Hubbard model on the Bethe lattice within DMFT using w2dynamics. 
In order to showcase the capabilities of \NAME to address low-temperature problems, we compare the continuous-time Quantum Monte Carlo (CTQMC) solver using the hybridization expansion method (CT-HYB) included in w2dynamics against the \NAME solver at finite temperature.

@@

\begin{figure}%[t!]
  \includegraphics[width=\linewidth]{figures/figBetheW2D.pdf}
    \caption{\label{figEx1W}%
    \textbf{Finite temperature DMFT solution.}
    Comparison of the imaginary parts of the Green's function $\Im{G}(i\omega_n)$ and Self-energy $\Im{\Sigma}(i\omega_n)$ from different solutions of the Hubbard model on the Bethe lattice using DMFT. Data are for $T/D=0.01$ and $U/D=2.0$. CTQMC results from the solver included with w2dynamics are compared against \NAME ED results, used both through the w2dynamics interface and with a standalone Fortran program. 
        }
\end{figure}

@@

\item Update the Weiss fields components, respectively 
  $\GG_0^{-1}$ and $\FF_0^{-1}$, through the {\bf self-con\-sis\-ten\-cy}
  relation: $\mathbfcal{G}^{-1}_0(i\omega) = {\mathbf G}_\mathrm{loc}(i\omega) +
  {\mathbf \Sigma}(i\omega)$ in Nambu space.

\item Optimize the bath parameters $\vec{x}$ to best describe the updated
    Weiss fields, potentially using \NAME provided conjugate gradient  fit
    procedures.
  \end{itemize}
%
The corresponding implementation in Python reads:
\begin{lstlisting}[style=mypython,numbers=none,basicstyle={\scriptsize\ttfamily}]
#DMFT CYCLE
converged=False;iloop=0
while (not converged and iloop<ed.Nloop ):
    iloop=iloop+1
    #Solve impurity problem
    ed.solve(bath)    
    #Self-consistency with aux_funx procedures:
    Smats = np.array([ed.get_sigma(axis="m",typ="n"),ed.get_sigma(axis="m",typ="a")])    
    Gmats = get_gloc(wm*1j       ,ed.xmu,HkNambu,Smats,axis="m") 
    Weiss = dmft_weiss_field(Gmats,Smats)                            
    #Fit
    bath = ed.chi2_fitgf(Weiss[0],Weiss[1],bath)
    #Error check
    err,converged=ed.check_convergence(Weiss,ed.dmft_error)
ed.finalize_solver()
\end{lstlisting}

@@

So we consider the model introduced in \secu{SecExamplesBetheDMFT}
with $U=0$ and the additional phononic and electron-phonon terms:
\begin{equation} \label{eqex:H_Holstein}
    H_\mathrm{int} = \sum_i \Big[\omega_0 b^\dagger_i b_i + g(b^\dagger_i +
    b_i)\sum_{\sigma}(c^\dagger_{i\sigma}c_{i\sigma}
    -\frac{1}{2})\Big]. 
\end{equation}
We consider the half-filling regime of the particle-hole symmetric Bethe lattice density of states. We set the half-bandwidth as our energy unit $D=1$ and introduced the electron-phonon coupling $\lambda = \tfrac{2g^2}{\omega_0D}$.  
The iterative DMFT solution algorithm follows the same principles
illustrated in the previous sections.

@@

In the anti-adiabatic limit, the Holstein interaction takes a
particularly simple form:
\begin{equation}\label{HlikeAttraction}
    H_\mathrm{int} \overset{ \omega_0 \rightarrow \infty}{ \longrightarrow } -\frac{\lambda}{2} \sum_i \Big[\sum_\sigma(c^\dagger_{i\sigma}c_{i\sigma} -\frac{1}{2}) \Big]^2
\end{equation}
which describes a local Hubbard-like attraction among electrons
(mediated by local phonons).
In the opposite, adiabatic, limit, $\omega_0 \rightarrow 0$ the system
enters in a Bipolaronic Insulating phase~\cite{Capone2006PRB} for our
choice of the coupling.

We characterize the model solution by showin the evolution of the
double occupation $\langle n_{\up}n_{\dw}\rangle$ as a function of the
phonon frequency $\omega_0$, see panel (A). In this panel we compare
the behavior for the normal solution ({\tt ed\_mode=normal}) against the superconductive
solution ({\tt ed\_mode=superc}). Our results capture the whole
crossover from adiabatic to anti-adiabatic regime. In the former
regime, the double occupation takes a similar value for the two
phases. However, approaching the anti-adiabatic regime, the two
solutions reach the limiting values corresponding to the residual
attraction \equ{HlikeAttraction}.   
To better characterize the nature of the superconducting
phase in the Holstein model we report the evolution of the anomalous order
parameter $\phi$ in the adiabatic to anti-adiabatic crossover.
The DMFT results obtained with \NAME show a rapid increase in the
superconducting order parameter as phonon frequency grows
large. Finally, in the anti-adiabatic regime, $\phi$ saturates to a
finite value corresponding to the attractive Hubbard-like interaction
of strength $\lambda/2$.

\begin{figure}[t!]
    \includegraphics[width=0.5\linewidth]
        {edipack2_examples/edipack2triqs/G_iw.pdf}
    \includegraphics[width=0.5\linewidth]
        {edipack2_examples/edipack2triqs/A_w.pdf}
    \caption{\label{figEx3}%
        Imaginary part of the Matsubara Green's function (left) and the
        corresponding orbital-resolved spectral function (right) computed for a
        three-orbital impurity model with an interaction of the Hubbard-Kanamori
        type (\ref{Hint}). This illustration is produced by the edipack2triqs
        example script presented in \secu{SecExamplesTRIQS}.
    }
\end{figure}

@@

We proceed to demonstrate how to use the edipack2triqs compatibility
layer (\secu{sSecInteropTRIQS}) to solve a quantum system comprised by a
3-orbital correlated impurity coupled to a few non-interacting bath sites.
The interaction term of the impurity Hamiltonian is in the Hubbard-Kanamori
form (\equ{Hint}).

A Python script implementing a calculation that makes use of edipack2triqs
generally begins with a few module imports.
\lstinputlisting[style=mypython,
                 numbers=none,
                 basicstyle={\scriptsize\ttfamily},
                 firstline=1, lastline=13]
{edipack2_examples/edipack2triqs/hubbard_kanamori.py}

One then proceeds to defining the system under consideration. In this case we
consider an impurity atom with three correlated orbitals and two bath states
per each impurity orbital. This information must be encoded in the fundamental
set objects that are later used to construct the solver.
\lstinputlisting[style=mypython,
                 numbers=none,
                 basicstyle={\scriptsize\ttfamily},
                 firstline=15, lastline=26]
{edipack2_examples/edipack2triqs/hubbard_kanamori.py}
The next step is to define a TRIQS many-body operator expression that represents
the Hamiltonian to be diagonalized.
\lstinputlisting[style=mypython,
                 numbers=none,
                 basicstyle={\scriptsize\ttfamily},
                 firstline=28, lastline=67]
{edipack2_examples/edipack2triqs/hubbard_kanamori.py}
Finally, one creates a solver object and performs the actual calculation by
calling its method {\tt solve()}. This is normally the most time- and
memory-consuming step.
\lstinputlisting[style=mypython,
                 numbers=none,
                 basicstyle={\scriptsize\ttfamily},
                 firstline=69, lastline=83]
{edipack2_examples/edipack2triqs/hubbard_kanamori.py}
The computation results are readily available as attributes of the solver
object. The following code snippet shows how to access measured expectation
values of the static observables, and how to employ the plotting framework of
TRIQS to visualize the obtained Matsubara and real-frequency impurity Green's
functions (\figu{figEx3}).
\lstinputlisting[style=mypython,
                 mathescape=false,
                 numbers=none,
                 basicstyle={\scriptsize\ttfamily},
                 firstline=85, lastline=112]
{edipack2_examples/edipack2triqs/hubbard_kanamori.py}

\subsection{Interacting Bernevig-Hughes-Zhang model (Fortran API, {\tt ed\_mode=nonsu2})}
In this section we focus on the DMFT solution of the interacting
Bernevig-Hughes-Zhang (BHZ).
As shown in Ref.~\onlinecite{AmaricciPRB2022}, at moderate
interactions this model admits an excitonic phase in which opposite
orbitals electrons and holes across the band gap bound into a coherent phase.

We consider a system of two-orbital electrons on a square
lattice interacting through a Hubbard-Kanamori term and describing
a quantum spin Hall insulator.
We consider a suitable matrix basis in terms of the Dirac
matrices $\Gamma_{a\a}=\sigma_a\otimes \tau_\a$, where $\sigma_a$ and
$\tau_\a$ are Pauli matrices, respectively, in the spin and orbital
pseudo-spin space. The  model Hamiltonian reads
$$
H = \sum_{k}\psi_{k}^\dagger H(k)\psi_{k} + H_{\rm int}
$$
where $\psi_{k}=[c_{1\uparrow k}, c_{2\uparrow k},
c_{1\downarrow k}, c_{2\downarrow k} ]$ is the spinor collecting
annihilation operator $c_{a\sigma k}$ destructing an electrons at
orbital $a=1,2$ with spin  $\sigma=\up,\dw$ and lattice momentum
$k$. The non-interacting part of the Hamiltonian is:
$$
H(k) = \left[M-2t(\cos{k_x}+\cos{k_y}) \right]\Gamma_{03} +
   \lambda\sin{k_x}\Gamma_{31} -   \lambda\sin{k_y}\Gamma_{02}
$$
where $M$ is the mass term, which plays the role of a crystal
field splitting among the orbitals. The presence of this term breaks
the symmetry in the orbital pseudo-spin channel.
The  interaction reads: 
$$
   H_{\rm int} = (U-J)\frac{\hat{N}(\hat{N}-1)}{2} - J\left( \frac{1}{4}\hat{N}^2 +
   \hat{S_z}^2 - 2 \hat{T_z}^2\right)
 $$
 where $\hat{N}=\tfrac{1}{2}\psi_i^\dagger \Gamma_{00}\psi_i$ is the
total density operator,
$\hat{S_z}=\tfrac{1}{2}\psi_i^\dagger \Gamma_{30}\psi_i$ is the total
spin polarization operator and $\hat{T_z}=\tfrac{1}{2}\psi_i^\dagger
\Gamma_{03}\psi_i$ is the orbital pseudo-spin polarization operator.
This form corresponds to the density-density part of the
Kanamori interaction. We neglect the pair-hopping and spin-flip purely
for numerical reasons.

@@

The code implementation follows the same guidelines discusses above
for the other examples. There are two specific modifications requied
to the  user. The first is to generate the Hamiltonian $H(k)$ on a discretized Brillouin zone in
order to obtain the local interacting Green's function, used in the
implementation of the self-consistency conditions:
$\GG^{-1}=G_\mathrm{loc}^{-1}+\Sigma$, with: 
$$
G_\mathrm{loc}(z) = \tfrac{1}{N_k}\sum_k \left[z+\mu-H(k)-\Sigma(z)
\right]^{-1}, 
$$
where $\z\in\CCC$  and $\Sigma(z)$ is the self-energy function obtained from the DMFT
solution of the proble.  $N_k$ is the number of $k$-points. 
The second is to  construct the ``renormalized'' topological
Hamiltonian
$$
H_\mathrm{top} = \sqrt{Z}[H(k) + \Re\Sigma(\omega\to0)]\sqrt{Z}, 
$$
where $Z$ is the renormalization constant. $H_\mathrm{top}$ describes the
low-energy properties of many-body problem and can be used to obtain a 
topological characterization of the interacting system.

@@

\paragraph{Results}
To capture the possible symmetry breaking in any excitonic channel
driven by the local interaction, we consider the vector order
parameter
$\vec{E}=[E_0,E_x,E_y,E_z]$ where $E_a=\langle\psi^\dagger
\Gamma_{a1}\psi \rangle$.
The first component $E_0$ describes the singlet excitonic state,
whereas the remaing ones correspond to the triplet states with
different spin orientation.   
The analysis of the strong-coupling regime as well as impurity
excitonic susceptibility available in \NAME suggest the possible
instability to an in-plane spin-triplet exciton phase, i.e. with $E_x$
and/or $E_y$ different from zero.
Interestingly, this state breakes several symmetries including
time-reversal and spin SU(2) which protect the topological state.

Thus, here we illustrate the DMFT solution of the excitonic phase as a way to
illustrate the ability of \NAME to capture matter phases with lowered
spin-symmetry ({\tt ed\_mode=nonsu2}) and the use of {\tt
  bath\_type=replica}.
\begin{lstlisting}[style=fstyle,numbers=none,basicstyle={\scriptsize\ttfamily}]
   !> Set $H_\mathrm{loc}$
   allocate(Hloc(Nso,Nso))
   Hloc = sum(Hk,dim=3)/Lk
   where(abs(dreal(Hloc))<1d-6)Hloc=zero
   call ed_set_hloc(Hloc)
   !> Setup the replica bath basis:
   allocate(lambdasym_vector(Nbath,4))
   allocate(Hsym_basis(Nso,Nso,4))
   Hsym_basis(:,:,1)=Gamma03  ;lambdasym_vector(:,1)= Mh
   Hsym_basis(:,:,2)=Gamma01 ;lambdasym_vector(:,2)= sb_field
   Hsym_basis(:,:,3)=Gamma31 ;lambdasym_vector(:,3)= sb_field
   Hsym_basis(:,:,4)=Gamma11 ;lambdasym_vector(:,4)=-sb_field
   !> Set the replica bath
   call ed_set_Hreplica(Hsym_basis,lambdasym_vector)
   !> Get bath dimension and allocate user bath to this size
   Nb=ed_get_bath_dimension(4)   !There are 4 symmetries
   allocate(Bath(Nb))
   !
   !> Initialize the ED solver (bath is guessed or read from file) 
   call ed_init_solver(bath)
\end{lstlisting}
where we use the local non-interacting Hamiltonian 
$H_\mathrm{loc}=\tfrac{1}{N_k}\sum_k H(k)$ to set 
$h^0_{\a\b\s\s'}$, i.e. the impurity Hamiltonian. To anticipate the
possibility of forming an excitonic ordered phase, the replica bath
is constructed out of matrix basis with 4 distinct elements
$\Gamma_{03}$, $\Gamma_{01}$, $\Gamma_{31}$, $\Gamma_{11}$ which
are proportional, respectively, to the mass term, the exciton singlet
$E_0$, the exciton triplet along easy-axis $E_z$ and in-plane
$E_x$ (we assume $E_y=0$ levaraging on residual $U(1)$ in-plane
symmetry. 
%
In the following we consider the case of $M=1$ which corresponds to a
QSHI in the non-interacting limit and $J/U=0.25$.
The implemention is nearly identical to the cases discussed above and
we report it here for completeness:
\begin{lstlisting}[style=fstyle,numbers=none,basicstyle={\scriptsize\ttfamily}]
   iloop=0;converged=.false.
   do while(.not.converged.AND.iloop<nloop)
     iloop=iloop+1
     !> Solve the impurity problem, retrieve $\Sigma(i\omega)$
     call ed_solve(bath)
     call ed_get_sigma(Smats,axis='mats')
     !> Get $G_\mathrm{loc}$ using $\color{comment-color}\mathrm{DMFTtools}$
     call get_gloc(Hk,Gmats,Smats,axis='m')
     !> Update the Weiss field ${\color{comment-color}\GG^{-1}=G^{-1}_\mathrm{loc}+\Sigma}$ +  Linear mixing: using $\color{comment-color}\mathrm{DMFTtools}$
     call dmft_self_consistency(Gmats,Smats,Weiss)
     if(iloop>1)Weiss = wmixing*Weiss + (1.d0-wmixing)*Weiss_;Weiss_=Weiss
     !> Fit to updated the bath
     call ed_chi2_fitgf(Weiss,bath,ispin=1)
     !Check convergence: using $\color{comment-color}\mathrm{DMFTtools}$
     converged = check_convergence(Weiss(1,1,:),dmft_error,nsuccess,nloop)
   enddo  
\end{lstlisting}

>>>>

The system is defined on a Bethe lattice with density of states
$\rho(\e)=\frac{1}{2D}\sqrt{D^2-\e^2}$,
where $D=2t$ is the half-bandwidth.
In DMFT, this lattice model is mapped onto a quantum impurity problem with
an effective electronic bath that must be determined
self-consistently.

Below, we present the key components of a basic DMFT implementation 
using \NAME for the Bethe lattice, employing either the Fortran or 
C++ APIs. Fully operational versions of the two codes can be found 
in the {\tt examples} directory of the \NAME source code. Both 
samples share a similar initial structure, including memory 
allocation, Bethe lattice creation, and solver initialization.

@@

\item[{\tiny {\it User}}] Update the Weiss field via the {\bf self-consistency}
  relation: $\GG^{-1}_0(i\omega) = G_\mathrm{loc}(i\omega) +
    \Sigma(i\omega)$. For the Bethe lattice, this simplifies to
    $\Delta = \tfrac{D}{4}G_\mathrm{loc}$.

@@

\subsubsection{Finite temperature solution (w2dynamics interface, CT-HYB benchmark)}\label{SecExamplesBetheDMFTw2d}
To demonstrate the usage of the w2dynamics interface to \NAME,  here we briefly discuss how to solve the same problem, i.e. the Hubbard model on the Bethe lattice within DMFT using w2dynamics. 
In order to showcase the capabilities of \NAME to address low-temperature problems, we compare the continuous-time Quantum Monte Carlo (CTQMC) solver using the hybridization expansion method (CT-HYB) included in w2dynamics against the \NAME solver at finite temperature.

@@

\begin{figure}%[t!]
  \includegraphics[width=\linewidth]{figures/figBetheW2D.pdf}
    \caption{\label{figEx1W}%
    \textbf{Finite temperature DMFT solution.}
    Comparison of the imaginary parts of the Green's function $\Im{G}(i\omega_n)$ and self-energy $\Im{\Sigma}(i\omega_n)$ from different solutions of the Hubbard model on the Bethe lattice using DMFT. Data are for $T/D=0.01$ and $U/D=2.0$. CTQMC results from the solver included with w2dynamics are compared against \NAME ED results, used both through the w2dynamics interface and with a standalone Fortran program. 
        }
\end{figure}

@@

\item Update the Weiss field's components, respectively 
  $\GG_0^{-1}$ and $\FF_0^{-1}$, through the {\bf self-con\-sis\-ten\-cy}
  relation: $\mathbfcal{G}^{-1}_0(i\omega) = {\mathbf G}_\mathrm{loc}(i\omega) +
  {\mathbf \Sigma}(i\omega)$ in Nambu space.

\item Optimize the bath parameters $\vec{x}$ to best describe the updated
    Weiss fields, potentially using the \NAME provided conjugate gradient  fit
    procedures.
  \end{itemize}
%
The corresponding implementation in Python reads:
\begin{lstlisting}[style=mypython,numbers=none,basicstyle={\scriptsize\ttfamily}]
#DMFT CYCLE
converged=False;iloop=0
while (not converged and iloop<ed.Nloop):
    iloop=iloop+1
    #Solve impurity problem
    ed.solve(bath)    
    #Self-consistency with aux_funx procedures:
    Smats = np.array([ed.get_sigma(axis="m",typ="n"),ed.get_sigma(axis="m",typ="a")])    
    Gmats = get_gloc(wm*1j       ,ed.xmu,HkNambu,Smats,axis="m") 
    Weiss = dmft_weiss_field(Gmats,Smats)                            
    #Fit
    bath = ed.chi2_fitgf(Weiss[0],Weiss[1],bath)
    #Error check
    err,converged=ed.check_convergence(Weiss,ed.dmft_error)
ed.finalize_solver()
\end{lstlisting}

@@

So we consider the model introduced in \secu{SecExamplesBetheDMFT}
with $U=0$ and the additional phononic and electron-phonon terms:
\begin{equation} \label{eqex:H_Holstein}
    H_\mathrm{int} = \sum_i \Big[\omega_0 b^\dagger_i b_i + g(b^\dagger_i +
    b_i)\sum_{\sigma}\left(c^\dagger_{i\sigma}c_{i\sigma}
    -\frac{1}{2}\right)\Big]. 
\end{equation}
We consider the half-filling regime of the particle-hole symmetric Bethe lattice density of states. We set the half-bandwidth as our energy unit $D=1$ and introduced the electron-phonon coupling $\lambda = \tfrac{2g^2}{\omega_0D}$.  
The iterative DMFT solution algorithm follows the same principles
illustrated in the previous sections.

@@

In the anti-adiabatic limit, the Holstein interaction takes a
particularly simple form:
\begin{equation}\label{HlikeAttraction}
    H_\mathrm{int} \overset{ \omega_0 \rightarrow \infty}{ \longrightarrow } -\frac{\lambda}{2} \sum_i \Big[\sum_\sigma\left(c^\dagger_{i\sigma}c_{i\sigma} -\frac{1}{2}\right) \Big]^2,
\end{equation}
which describes a local Hubbard-like attraction among electrons
(mediated by local phonons).
In the opposite, adiabatic, limit, $\omega_0 \rightarrow 0$ the system
enters a Bipolaronic Insulating phase~\cite{Capone2006PRB} for our
choice of the coupling.

We characterize the model solution by showing the evolution of the
double occupation $\langle n_{\up}n_{\dw}\rangle$ as a function of the
phonon frequency $\omega_0$, see panel (A) of \figu{figEx5}. In this panel we compare
the behavior for the normal solution ({\tt ed\_mode=normal}) against the superconductive
solution ({\tt ed\_mode=superc}). Our results capture the whole
crossover from adiabatic to anti-adiabatic regime. In the former
regime, the double occupation takes a similar value for the two
phases. However, approaching the anti-adiabatic regime, the two
solutions reach the limiting values corresponding to the residual
attraction \equ{HlikeAttraction}.   
To better characterize the nature of the superconducting
phase in the Holstein model, we report the evolution of the anomalous order
parameter $\phi$ in the adiabatic to anti-adiabatic crossover (panel (B) of \figu{figEx5}).
The DMFT results obtained with \NAME show a rapid increase in the
superconducting order parameter as phonon frequency grows
large. Finally, in the anti-adiabatic regime, $\phi$ saturates to a
finite value corresponding to the attractive Hubbard-like interaction
of strength $\lambda/2$.

\begin{figure}[t!]
    \includegraphics[width=0.5\linewidth]
        {edipack2_examples/edipack2triqs/G_iw.pdf}
    \includegraphics[width=0.5\linewidth]
        {edipack2_examples/edipack2triqs/A_w.pdf}
    \caption{\label{figEx3}%
        Imaginary part of the Matsubara Green's function $G_\alpha(i\omega_n)$ (left) and the
        corresponding orbital-resolved spectral function $A_\alpha(\omega) = -\Im{G_\alpha(\omega)} / \pi$ (right) computed for a
        three-orbital impurity model with an interaction of the Hubbard-Kanamori
        type (\ref{Hint}). This illustration is produced by the EDIpack2TRIQS
        example script presented in \secu{SecExamplesTRIQS}.
    }
\end{figure}

@@

We proceed to demonstrate how to use the EDIpack2TRIQS compatibility
layer (\secu{sSecInteropTRIQS}) to solve a quantum system comprised by a
3-orbital correlated impurity coupled to a few non-interacting bath sites.
The interaction term of the impurity Hamiltonian is in the Hubbard-Kanamori
form (\equ{Hint}).

A Python script implementing a calculation that makes use of EDIpack2TRIQS
generally begins with a few module imports.
\lstinputlisting[style=mypython,
                 numbers=none,
                 basicstyle={\scriptsize\ttfamily},
                 firstline=1, lastline=13]
{edipack2_examples/edipack2triqs/hubbard_kanamori.py}

One then proceeds to defining the system under consideration. In this case we
consider an impurity atom with three correlated orbitals and two bath states
per each impurity orbital. This information must be encoded in the fundamental
set objects that are later used to construct the solver.
\lstinputlisting[style=mypython,
                 numbers=none,
                 basicstyle={\scriptsize\ttfamily},
                 firstline=15, lastline=26]
{edipack2_examples/edipack2triqs/hubbard_kanamori.py}
The next step is to define a TRIQS many-body operator expression that represents
the Hamiltonian to be diagonalized.
\lstinputlisting[style=mypython,
                 numbers=none,
                 basicstyle={\scriptsize\ttfamily},
                 firstline=28, lastline=67]
{edipack2_examples/edipack2triqs/hubbard_kanamori.py}
Finally, one creates a solver object and performs the actual calculation by
calling its method {\tt solve()}. This is normally the most time- and
memory-consuming step.
\lstinputlisting[style=mypython,
                 numbers=none,
                 basicstyle={\scriptsize\ttfamily},
                 firstline=69, lastline=83]
{edipack2_examples/edipack2triqs/hubbard_kanamori.py}
The computation results are readily available as attributes of the solver
object. The following code snippet shows how to access measured expectation
values of the static observables, and how to employ the plotting framework of
TRIQS to visualize the obtained Matsubara and real-frequency impurity Green's
functions (\figu{figEx3}).
\lstinputlisting[style=mypython,
                 mathescape=false,
                 numbers=none,
                 basicstyle={\scriptsize\ttfamily},
                 firstline=85, lastline=113]
{edipack2_examples/edipack2triqs/hubbard_kanamori.py}

\subsection{Interacting Bernevig-Hughes-Zhang model (Fortran API, {\tt ed\_mode=nonsu2})}
In this section we focus on the DMFT solution of the interacting
Bernevig-Hughes-Zhang (BHZ).
As shown in Ref.~\onlinecite{AmaricciPRB2022}, at moderate
interactions this model admits an excitonic phase in which opposite
orbital electrons and holes across the band gap bind into a coherent phase.

We consider a system of two-orbital electrons on a square
lattice interacting through a Hubbard-Kanamori term and describing
a quantum spin Hall insulator.
We consider a suitable matrix basis in terms of the Dirac
matrices $\Gamma_{a\a}=\sigma_a\otimes \tau_\a$, where $\sigma_a$ and
$\tau_\a$ are Pauli matrices, respectively, in the spin and orbital
pseudo-spin space. The  model Hamiltonian reads
$$
H = \sum_{k}\psi_{k}^\dagger H(k)\psi_{k} + H_{\rm int},
$$
where $\psi_{k}=[c_{1\uparrow k}, c_{2\uparrow k},
c_{1\downarrow k}, c_{2\downarrow k} ]$ is the spinor collecting
annihilation operator $c_{a\sigma k}$ destroying an electron at
orbital $a=1,2$ with spin  $\sigma=\up,\dw$ and lattice momentum
$k$. The non-interacting part of the Hamiltonian is:
$$
H(k) = \left[M-2t(\cos{k_x}+\cos{k_y}) \right]\Gamma_{03} +
   \lambda\sin{k_x}\Gamma_{31} -   \lambda\sin{k_y}\Gamma_{02},
$$
where $M$ is the mass term, which plays the role of a crystal
field splitting among the orbitals. The presence of this term breaks
the symmetry in the orbital pseudo-spin channel.
The  interaction reads: 
$$
   H_{\rm int} = (U-J)\frac{\hat{N}(\hat{N}-1)}{2} - J\left( \frac{1}{4}\hat{N}^2 +
   \hat{S_z}^2 - 2 \hat{T_z}^2\right),
 $$
 where $\hat{N}=\tfrac{1}{2}\psi_i^\dagger \Gamma_{00}\psi_i$ is the
total density operator,
$\hat{S_z}=\tfrac{1}{2}\psi_i^\dagger \Gamma_{30}\psi_i$ is the total
spin polarization operator and $\hat{T_z}=\tfrac{1}{2}\psi_i^\dagger
\Gamma_{03}\psi_i$ is the orbital pseudo-spin polarization operator.
This form corresponds to the density-density part of the
Kanamori interaction. We neglect the pair-hopping and spin-flip purely
for numerical reasons.

@@

The code implementation follows the same guidelines discussed above
for the other examples. There are two specific modifications required
from the user. The first is to generate the Hamiltonian $H(k)$ on a discretized Brillouin zone in
order to obtain the local interacting Green's function, used in the
implementation of the self-consistency conditions:
$\GG^{-1}=G_\mathrm{loc}^{-1}+\Sigma$, with 
$$
G_\mathrm{loc}(z) = \tfrac{1}{N_k}\sum_k \left[z+\mu-H(k)-\Sigma(z)
\right]^{-1}, 
$$
where $\z\in\CCC$  and $\Sigma(z)$ is the self-energy function obtained from the DMFT
solution of the proble.  $N_k$ is the number of $k$-points. 
The second is to  construct the ``renormalized'' topological
Hamiltonian
$$
H_\mathrm{top} = \sqrt{Z}[H(k) + \Re\Sigma(\omega\to0)]\sqrt{Z}, 
$$
where $Z$ is the renormalization constant. $H_\mathrm{top}$ describes the
low-energy properties of many-body problem and can be used to obtain a 
topological characterization of the interacting system.

@@

\paragraph{Results}
To capture the possible symmetry breaking in any excitonic channel
driven by the local interaction, we consider the vector order
parameter
$\vec{E}=[E_0,E_x,E_y,E_z]$, where $E_a=\langle\psi^\dagger
\Gamma_{a1}\psi \rangle$.
The first component $E_0$ describes the singlet excitonic state,
whereas the remaining ones correspond to the triplet states with
different spin orientation.   
The analysis of the strong-coupling regime as well as impurity
excitonic susceptibility available in \NAME suggest the possible
instability to an in-plane spin-triplet exciton phase, i.e. with $E_x$
and/or $E_y$ different from zero.
Interestingly, this state breaks several symmetries including
time-reversal and spin SU(2), which protect the topological state.

Thus, here we illustrate the DMFT solution of the excitonic phase as a way to
illustrate the ability of \NAME to capture matter phases with lowered
spin-symmetry ({\tt ed\_mode=nonsu2}) and the use of {\tt
  bath\_type=replica}.
\begin{lstlisting}[style=fstyle,numbers=none,basicstyle={\scriptsize\ttfamily}]
   !> Set $H_\mathrm{loc}$
   allocate(Hloc(Nso,Nso))
   Hloc = sum(Hk,dim=3)/Lk
   where(abs(dreal(Hloc))<1d-6)Hloc=zero
   call ed_set_hloc(Hloc)
   !> Setup the replica bath basis:
   allocate(lambdasym_vector(Nbath,4))
   allocate(Hsym_basis(Nso,Nso,4))
   Hsym_basis(:,:,1)=Gamma03  ;lambdasym_vector(:,1)= Mh
   Hsym_basis(:,:,2)=Gamma01 ;lambdasym_vector(:,2)= sb_field
   Hsym_basis(:,:,3)=Gamma31 ;lambdasym_vector(:,3)= sb_field
   Hsym_basis(:,:,4)=Gamma11 ;lambdasym_vector(:,4)=-sb_field
   !> Set the replica bath
   call ed_set_Hreplica(Hsym_basis,lambdasym_vector)
   !> Get bath dimension and allocate user bath to this size
   Nb=ed_get_bath_dimension(4)   !There are 4 symmetries
   allocate(Bath(Nb))
   !
   !> Initialize the ED solver (bath is guessed or read from file) 
   call ed_init_solver(bath)
\end{lstlisting}
Here we use the local non-interacting Hamiltonian 
$H_\mathrm{loc}=\tfrac{1}{N_k}\sum_k H(k)$ to set 
$h^0_{\a\b\s\s'}$, i.e. the impurity Hamiltonian. To anticipate the
possibility of forming an excitonic ordered phase, the replica bath
is constructed out of the matrix basis with 4 distinct elements
$\Gamma_{03}$, $\Gamma_{01}$, $\Gamma_{31}$, $\Gamma_{11}$ which
are proportional, respectively, to the mass term, the exciton singlet
$E_0$, the exciton triplet along easy-axis $E_z$ and in-plane
$E_x$ (we assume $E_y=0$ leveraging on residual $U(1)$ in-plane
symmetry). 
%
In the following we consider the case of $M=1$, which corresponds to a
QSHI in the non-interacting limit and $J/U=0.25$.
The implementation is nearly identical to the cases discussed above and
we report it here for completeness:
\begin{lstlisting}[style=fstyle,numbers=none,basicstyle={\scriptsize\ttfamily}]
   iloop=0;converged=.false.
   do while(.not.converged.AND.iloop<nloop)
     iloop=iloop+1
     !> Solve the impurity problem, retrieve $\Sigma(i\omega)$
     call ed_solve(bath)
     call ed_get_sigma(Smats,axis='mats')
     !> Get $G_\mathrm{loc}$ using $\color{comment-color}\mathrm{DMFTtools}$
     call get_gloc(Hk,Gmats,Smats,axis='m')
     !> Update the Weiss field ${\color{comment-color}\GG^{-1}=G^{-1}_\mathrm{loc}+\Sigma}$ +  Linear mixing: using $\color{comment-color}\mathrm{DMFTtools}$
     call dmft_self_consistency(Gmats,Smats,Weiss)
     if(iloop>1)Weiss = wmixing*Weiss + (1.d0-wmixing)*Weiss_;Weiss_=Weiss
     !> Fit to updated the bath
     call ed_chi2_fitgf(Weiss,bath,ispin=1)
     !Check convergence: using $\color{comment-color}\mathrm{DMFTtools}$
     converged = check_convergence(Weiss(1,1,:),dmft_error,nsuccess,nloop)
   enddo  
\end{lstlisting}

%
The main effect of the interaction on the topological properties can be captured by the renormalized mass term $M_\mathrm{eff}=M+\tfrac{1}{4}\Tr{\Re\Sigma(i\omega\to0)}$. The
real part of the self-energy being proportional to $\Gamma_{03}$
corrects the mass term with respect to its bare value. To leading
order this correction is proportional to $\tfrac{U-5J}{4}\langle
\hat{T}_z\rangle$. So, for our choice of parameters, the effect of
interaction would be to reduce the (effective) mass term and thus
reducing the topological band gap. In the strong-coupling limit the
two orbitals get populated by one electron per site, reaching the
conditions for the formation of a Mott insulating state.
As reported in panel (A) of \figu{figEx4}, at intermediate coupling,
the DMFT solution features the formation of a
region of exciton condensation with $\langle E_x\rangle>0$, $\langle
E_0\rangle=\langle E_z\rangle=0$. 
Unlike the static mean-field description~\cite{Blason}, the
transition from the QSHI topological state to the Excitonic Insulator 
(EI) becomes of first-order in the presence of the dynamical fluctuations~\cite{Continentino,Paoletti,BellomiaKMH} contained in DMFT. On the
other hand, the EI continuously evolves into a Mott Insulator (MI)
for larger interaction (neglecting for simplicity any 
antiferromagnetic ordering that would naturally occur in this
regime).
In the same panel, we report the evolution of
$\langle T_z\rangle$ to illustrate the progressive reduction of the
orbital polarization and, thus, of the effective mass.
